Data preparation B

### 2020-01-13 script/data-prep-B.Rmd
process 1999 data

---
title: 'Data preparation'
author: ""
output:
    html_document:
        smart: false
editor_options: 
  chunk_output_type: console
---

@@

@@

```{r }
library(sjlabelled)
library(dplyr)
library(haven)

@@

## Get data 
```{r results='hide'}

issp99<- read_stata("../input/data/origin data/dataissp/issp1999.dta")
issp09<- read_stata("../input/data/origin data/dataissp/issp2009.dta")
issp19<- read_stata("../input/data/origin data/dataissp/issp2019.dta")

@@

## Select Chile
```{r seleccionar Chile}

# 1999

sjmisc::find_var(issp99,"Chile",search = "all")
sjlabelled::get_labels(issp99$v3,values = "as.name")

cl99 = issp99 %>% dplyr::filter(v3==30)

# 2009

sjmisc::find_var(issp09,"Chile",search = "all")
sjlabelled::get_labels(issp09$V4,values = "as.name")

cl09 <- issp09 %>% dplyr::filter(V4==152)

# 2019 ... solo Chile

cl19 <- issp19

@@

## Lowercase & save

```{r}
bases=c("cl99","cl09","cl19")

for (i in bases) {
  df.tmp <- get(i)
  names(df.tmp) <- tolower(names(df.tmp)) 
  assign(i, df.tmp)
  save(list=paste(i, sep=""), file = (paste("../input/data/dataproc/", i,".RData", sep="")))
}

# database for analysis 1999 - 2009

* This dataset includes all 3 waves with salary gaps variables, goverment responsibility on inequality, and meritocracy perception and preferences

```{r}
library(dplyr)
```

```{r pregunta redistribucion}
load("../input/data/dataproc/cl99.RData")
load("../input/data/dataproc/cl09.RData")
load("../input/data/dataproc/cl19.RData")
```

```{r agno 1999}

dat99 <- cl99 %>% dplyr::select(v3,
                         v21,v16, # Salario percibio obrero y gerente
                         v31,v26, # Salario justo    obrero y gerente
                         v35,     # redistribucion
                         x_degr,# Educación 
                         v6, # perMeritocracia
                         rincome) #, ingreso
#--- missing meritocracia -------------------------------------------------#

#dat99$v6 <- car::recode(dat99$v6,"c(8,9)=NA",as.numeric = TRUE)
dat99$v6[dat99$v6==8]<-NA
dat99$v6[dat99$v6==9]<-NA

#dat99$rincome <- car::recode(dat99$rincome,"c(999997,999998,999999)=NA",as.numeric = TRUE)
dat99$rincome[dat99$rincome==999997]<-NA
dat99$rincome[dat99$rincome==999998]<-NA
dat99$rincome[dat99$rincome==999999]<-NA

# Redistribucion
#dat99$v35 <- car::recode(dat99$v35,"c(8,9)=NA",as.numeric = TRUE)
dat99$v35[dat99$v35==8]<-NA
dat99$v35[dat99$v35==9]<-NA

dat99$v35 <- sjmisc::rec(dat99$v35,rec="rev")
#Educacion
#dat99$x_degr <- car::recode(dat99$x_degr, "c(3009)=NA", as.numeric=TRUE)
dat99$x_degr[dat99$x_degr==3009]<-NA

# Rename variables
dat99 <- dplyr::rename(dat99,
               pais=v3,
               salperger=v16,
               salperobr=v21,
               saljusger=v26,
               saljusobr=v31,
               permer = v6,
               redis=v35,
               educ_rec=x_degr,
               income=rincome)
#--- missing brechas -------------------------------------------------#
names(dat99)
dat99[2:5][dat99[2:5] >= 99999999998] <- NA
dat99 <- dat99 %>%
  mutate(salperobr=replace(salperobr, salperobr <= 40000 | salperobr>= 1000000, NA)) %>%
  mutate(salperger=replace(salperger, salperger <= 250000 | salperger>= 100000001, NA)) %>%
  mutate(saljusobr=replace(saljusobr, saljusobr <= 40000, NA)) %>%
  mutate(saljusger=replace(saljusger, saljusger <= 100000, NA))
#---Brecha salarial percibida
dat99$gap_perc   <-  as.numeric(dat99$salperger/dat99$salperobr) # diferencia total
dat99$lngap_perc <-  as.numeric(log(dat99$gap_perc))                 # diferencia log
#---Brecha salarial justa
dat99$gap_just   <-  as.numeric(dat99$saljusger/dat99$saljusobr) # diferencia total
dat99$lngap_just <-  as.numeric(log(dat99$gap_just))                # diferencia log
dim(na.omit(dat99))

#---Recodificacion Educacion
#dat99 <- mutate(dat99, educ_rec = car::recode(dat99$educ, "3001:3004 = 1;  = 2;3008=2; 3010 = 3;3007 = 4"))  
dat99$educ_rec[dat99$educ_rec== 3001:3004 ] <- 1
dat99$educ_rec[dat99$educ_rec== 3005:3006 ] <- 2
dat99$educ_rec[dat99$educ_rec== 3008] <- 2
dat99$educ_rec[dat99$educ_rec== 3010] <- 3
dat99$educ_rec[dat99$educ_rec== 3007] <- 4

sjPlot::view_df(cl99, show.frq = T)
dat99 <- na.omit(dat99)
cor(dat99)

#redis

dat99$redis<-ifelse(dat99$redis>=8,1,0)
hist(dat99$redis)

dat99$universitaria<-ifelse(dat99$educ_rec==4,1,0)
dat99$tecnico<-ifelse(dat99$educ_rec==3,1,0)
dat99$media<-ifelse(dat99$educ_rec==2,1,0)
dat99$basica<-ifelse(dat99$educ_rec==1,1,0)

#permerit !!!Esta pregunta es distinta
dat99$permer<-ifelse(dat99$permer<=2,1,0)
hist(dat99$permer)

#dat99 <- mutate(dat99, income_rec = car::recode(dat99$income, "1:4 = 1; 5:6 = 2; 7:8=3; 9:10=4; 11=5; 12:14=6"))  
dat99$income_rec[dat99$income== 1:4] <- 1
dat99$income_rec[dat99$income== 5:6] <- 2
dat99$income_rec[dat99$income== 7:8] <- 3
dat99$income_rec[dat99$income== 9:10] <- 4
dat99$income_rec[dat99$income== 11] <- 5
dat99$income_rec[dat99$income== 12:14] <- 6

dat99$bajo <- ifelse(dat99$income_rec==1,1,0) 
dat99$mediobajo <- ifelse(dat99$income_rec==2,1,0)
dat99$medio <- ifelse(dat99$income_rec==3,1,0)
dat99$medioalto <- ifelse(dat99$income_rec==4,1,0)
dat99$alto <- ifelse(dat99$income_rec==5,1,0)
dat99$muyalto <- ifelse(dat99$income_rec==6,1,0)

```
__Condificacion de educación__

* Educacion Basica o menos: SE, EBI, EBC, EMI.
* Educación Media: EMC, ENUI,EUI.
* Educacion Tecnica: ENUC
* Educacion Univercitaria o más:  EUC, PG. (EN 1999 no hay pg)

>>>>

---
title: 'Data preparation B'
author: ""
output:
    html_document:
        smart: false
editor_options:
  chunk_output_type: console
---

@@

# Data preparation B: selects & recodes

@@

```{r libraries}
library(sjlabelled)
library(dplyr)
library(haven)
library(descr)
library(sjmisc)

@@

## Variables:

a- salaries perceived & just worker & manager
b- education
c- income
d- redistribution

## Year 1999

### Load data

```{r }
load("input/data/intermediate/cl99.RData") # cl99
```

### a- salaries perceived & just worker & manager

# Find vars

find_var(cl99,"worker",search = "all") # v21: perceived worker, v31 just worker unrecoded
find_var(cl99,"chairman",search = "all") # v16: perceived chairman, v26 just chairman unrecoded

# Create object with vars & rename

dat99 <- cl99 %>% select(v21,v31,v16,v26) %>% rename(salperger=v16,
               salperobr=v21,
               saljusger=v26,
               saljusobr=v31)

# Recodes

stargazer(dat99, type="text")

dat99[][dat99  >= 99999999998] <- NA # missings

dat99 <- dat99 %>%
  mutate(salperobr=replace(salperobr, salperobr <= 40000 | salperobr>= 1000000, NA)) %>%
  mutate(salperger=replace(salperger, salperger <= 250000 | salperger>= 100000001, NA)) %>%
  mutate(saljusobr=replace(saljusobr, saljusobr <= 40000, NA)) %>%
  mutate(saljusger=replace(saljusger, saljusger <= 100000, NA))

#---Brecha salarial percibida
dat99$gap_perc   <-  as.numeric(dat99$salperger/dat99$salperobr) # diferencia total
dat99$lngap_perc <-  as.numeric(log(dat99$gap_perc))                 # diferencia log

#---Brecha salarial justa
dat99$gap_just   <-  as.numeric(dat99$saljusger/dat99$saljusobr) # diferencia total
dat99$lngap_just <-  as.numeric(log(dat99$gap_just))                # diferencia log
dim(na.omit(dat99))
```

### b - education

```{r}

# Find vars

find_var(cl99,"degr",search = "all") # v21: perceived worker, v31 just worker
frq(as.factor(cl99$x_degr))
frq(as.factor(cl99$degree))

# Problema: no se encuentran los codigos originales en codebook, se derivan de la pregunta father education v317, que tiene las 7 categorías unesco asumiendo que las 7 de degree son las mismas:

# 1. None
# 2. Incomplete primary
# 3. Primary completed
# 4. Incomplete secondary (also includes completed qualifications at a level below academic secondary school
# completion – for example skilled manual qualifications given by trade schools.)
# 5. Secondary completed
# 6. Some tertiary education but less than a university Bachelor’s degree (Includes incomplete university. Also
# includes completed post-secondary courses below university Bachelor’s degree level. Examples are qualifications
# from US junior college, polytechnical institutes in many countries, and other similar post-secondary education.)
# 7. University completed (Bachelor’s degree or higher)

# Codificación original Francisco queda stand-by:

# #dat99 <- mutate(dat99, educ_rec = car::recode(dat99$educ, "3001:3004 = 1;  = 2;3008=2; 3010 = 3;3007 = 4"))
# dat99$educ_rec[dat99$educ_rec== 3001:3004 ] <- 1
# dat99$educ_rec[dat99$educ_rec== 3005:3006 ] <- 2
# dat99$educ_rec[dat99$educ_rec== 3008] <- 2
# dat99$educ_rec[dat99$educ_rec== 3010] <- 3
# dat99$educ_rec[dat99$educ_rec== 3007] <- 4

cl99 <- cl99 %>% select(degree) %>% mutate(educ_rec=rec(degree,rec = "1:4=1; 5=2 ; 6=3 ; 7=4"))

names(dat99)

dat99<-cbind(dat99,cl99 %>% select(educ_rec)) # add educ to table

summary(dat99)

dat99 %>% group_by(educ_rec) %>% summarize(m=median(gap_perc, na.rm = TRUE))
```

### c-income

```{r}

find_var(cl99,"income",search = "all") # v21: perceived worker, v31 just worker

freq(cl99$incomer) # deciles family
freq(cl99$rincomer) # deciles respondent, more missings
freq(cl99$income) # 14 categories family
freq(cl99$rincome) # categories respondent

# take income, impute middle range, then divide by persons in household, then deciles

cl99$income_c= rec(cl99$income,rec =   
      "1=45000;
      2=105500;
      3=135500;
      4=165500;
      5=195500;
      6=225500;
      7=265500;
      8=240500;
      9=495500;
      10=800500;
      11=1250000;
      12=1750000;
      13=2500000;
      14=4000000;
      999998=NA;
      999999=NA")

freq(cl99$income_c) # ok

find_var(cl99,"household",search = "all") # persons in household
freq(cl99$hompop)

dat99$ingreso_pc= cl99$income_c/cl99$hompop

# Generar quintiles de ingreso

dat99 %>% mutate(Quint = ntile(ingreso_pc,5)) -> dat99

# Check
freq( dat99$Quint)

@@

### d - redistribution

```{r}
find_var(cl99,"income",search = "all") # v35
freq(cl99$v35)
get_labels(cl99$v35,values = "as.name")

dat99$redis <-rec(cl99$v35, rec="8:9=NA;else=copy")
dat99$redis <-rec(dat99$redis, rec="rev")
freq(dat99$redis)
dat99$redis <-set_labels(dat99$redis,labels= c("Muy en desacuerdo"=1, "En descacuerdo"=2,"Ni acuerdo ni desacuerdo"=3, "De acuerdo"=4, "Muy de acuerdo"=5))
get_labels(dat99$redis, values="as.name")
freq(to_label(dat99$redis))
```

### Save 99 data

```{r}
save(dat99,file = "input/data/processed/dat99.RData")

@@

## Year 2009
